Estimating Predictive Rate-Distortion via Neural Variational Inference

### 2019-03-05 paper.tex
further analyses

@@

@@

\section{Conclusion}
We introduce Neural Predictive Rate Distortion (NPRD), a method for estimating predictive rate-distortion when only sample trajectories are given.
Unlike OCE, the most general prior method, NPRD scales to long sequences and large state spaces.
On analytically tractable processes, we show that it closely fits the analytical rate-distortion curve and recovers the causal states of the process.
On POS-level modeling of natural language, it agrees with OCE in the setting of low rate and short sequences; outside these settings, OCE fails due to combinatorial explosion and overfitting, while NPRD continues to provide estimates.
Finally, we use NPRD to provide the first estimates of predictive rate-distortion for word-level modeling of natural language.

>>>>

\begin{figure*}
\includegraphics[width=0.3\textwidth]{code/figures/en-words-info-fitted.pdf}
\includegraphics[width=0.3\textwidth]{code/figures/en-words-nlogbeta-mem-fitted.pdf}
\includegraphics[width=0.3\textwidth]{code/figures/en-words-nlogbeta-ee-fitted.pdf}
	\caption{Word-level modeling of English.}\label{fig:en-logbeta-fit}
\end{figure*}

\begin{figure*}
\includegraphics[width=0.3\textwidth]{code/figures/ru-words-info-fitted.pdf}
\includegraphics[width=0.3\textwidth]{code/figures/ru-words-nlogbeta-mem-fitted.pdf}
\includegraphics[width=0.3\textwidth]{code/figures/ru-words-nlogbeta-ee-fitted.pdf}
	\caption{Word-level modeling of English.}\label{fig:ru-logbeta-fit}
\end{figure*}

@@

\section{Conclusion}
We introduce Neural Predictive Rate Distortion (NPRD), a method for estimating predictive rate-distortion when only sample trajectories are given.
Unlike OCE, the most general prior method, NPRD scales to long sequences and large state spaces.
On analytically tractable processes, we show that it closely fits the analytical rate-distortion curve and recovers the causal states of the process.
On POS-level modeling of natural language, it agrees with OCE in the setting of low rate and short sequences; outside these settings, OCE fails due to combinatorial explosion and overfitting, while NPRD continues to provide estimates.
Finally, we use NPRD to provide the first estimates of predictive rate-distortion for modeling natural language.

@@
